Optimising Heterogeneous Ambulance Fleet Allocations in Jakarta

### 2024-07-26 response/response.tex

\documentclass{article}

\usepackage{fullpage}
\usepackage{parskip}
\usepackage{setspace}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{hyperref}

\title{Responses to Reviewers}
\author{}
\date{}

\begin{document}

\maketitle

We would like to express our gratitude to the editor, associate editor, and reviewers for facilitating the review process, and for their useful suggestions and remarks on the manuscript.

This document will address the specific concerns raised.

\section{Responses to the Editor}

\begin{itemize}

\item The editor wrote:
\begin{quote}
``Please cite relevant, recent work published in this journal. This is by no means mandatory, but it will help to position your work in the context of research published in the journal.''
\end{quote}

\item The editor wrote:
\begin{quote}
``The manuscript does not sufficiently discuss the managerial implications of the work. Please develop this further in your revision.''
\end{quote}

\end{itemize}

\section{Responses to the Associate Editor}
There were no comments by the associate editor. We would like to thank the associate editor for their work facilitating the review process.

\section{Responses to the Reviewers}

\subsection{Comments from Reviewer 1}

\begin{itemize}

\item Reviewer 1 wrote:
\begin{quote}
``No Verification \& Validation of the simulation model is described. At least some discussion regarding the validity of the model has to be included in the case study.''
\end{quote}

\item Reviewer 1 wrote:
\begin{quote}
``No measures of variance are given for the simulation results in Tables 1 and 4. Especially since repeated simulation results were conducted, measures of variance should be included to investigate the statistical significance of the results.''
\end{quote}

\item Reviewer 1 wrote:
\begin{quote}
``Algorithm 1: Updating the mutation rate m by subtracting the maximum of m0 and a second number from m0 will lead to a negative (or zero) mutation rate, which does not make sense. Probably should be min instead of max.''
\end{quote}
The algorithm is correct. Line 5 of the algorithm states that $m_0 \leftarrow \lceil m_0 - \max(m_0, i \times c) \rceil$. Without the ceiling function, this would have the effect of linearly decreasing wiht $i$ until 0, and then remaining at 0 thereafter. With the ceiling function, the linear part becomes a stepwise decent to 0, and then 0 thereafter.

\item Reviewer 1 wrote:
\begin{quote}
``Unclear, why the subchapter 4.1 on survival functions is necessary, since traditional coverage targets are used in the case study and the model formulation does not depend on characteristics of the survival functions. While survival functions are an important performance metric in the reviewers opinion, it is not clear why the reader needs this information to follow the remainder of the paper.''
\end{quote}
We disagree with the reviewer here. In the case study, traditional targets such a mean response times are reported alongside survival measures, based on the survival functions. However, a key part of the paper is the optimisation process, and all scenarios compared are `optimised' scenarios that result from this process. The objective function of the optimisation process maximises survival, based on survival functions. This is evnidenced in Equations 6 and 7, which are component parts of the objective function given in Equation 5. Therefore the survival functions are central to the work presented.

\item Reviewer 1 wrote:
\begin{quote}
``p. 8/9: A short explanation, of what the MINPACK hybrd and hybrj algorithms do and how they work should be included either directly or in the appendix.''
\end{quote}

\item Reviewer 1 wrote:
\begin{quote}
``There is no discussion of how well the proposed model estimates survival probabilities. This could be easily done by calculating the survival probabilites based on the simulated response times and comparing the simulated survival probabilities with the ones calculated by the MESLMHPHF. It is also not clear, why e.g. in Table 4 the expected Survival from the MESLMHPHF was used instead of the more realistic survival based on the simulation results.''
\end{quote}

\item Reviewer 1 wrote:
\begin{quote}
``Formulating the MESLMHPHF as an explicit optimization model with an objective function and constraints would make it easier to follow the model logic than it is with the separation of the formulation over multiple pages.''
\end{quote}
This has now been given as an Appendix, while the body of the paper presents the model in a way that explains and justifies each step.

\item Reviewer 1 wrote:
\begin{quote}
``p.1 l. 37 ‘driven by inadequate a lack of financial investment and human capital’ (delete inadequate)''
\end{quote}
Fixed.

\item Reviewer 1 wrote:
\begin{quote}
``p.5, l. 21: $J_{yk}$ was not defined, only $J_k$''
\end{quote}
Fixed, all instances should the $J_k$.

\item Reviewer 1 wrote:
\begin{quote}
``Fig. 10: Both plots should have the same y-axis interval and ideally start
at 0''
\end{quote}
The scale of Fig 10.b is so small that the lines would not be able to be differentiated if it had the same scale as Fig 10.a; and would be impossible to see at all if the y-axis starts at 0. Instead, we have included a sentence in the caption to note that the y-axes are different for both plots.

\item Reviewer 1 wrote:
\begin{quote}
``Fig. 11: A1 survival does not increase monotonically, especially in the single vehicle type scenario. A discussion of possible causes for this counterintuitive result should be included.''
\end{quote}
Thank you for this comment. We have now added some sentences explaining this phenomenon, calling back to the behaviour of the survival function itself and arguing that this is an upper bound for the survival of A1 patients.

\item Reviewer 1 wrote:
\begin{quote}
``Equations 8 and 9: It is stated, that $\beta_{p a_1 a_2}$ indicates, that a vehicle from a1 can reach p quicker, than a vehicle from a2. The equations 8 and 9 however use $\leq$ and not < , which is not consistent with the description.''
\end{quote}
Thank you for spotting this. As the $\beta_{p a_1 a_2}$ define an ordering of the ambulance locations, it does not matter whether they are defined by $\leq$ or <, as we assume that no two distinct ambulance locations are equidistant from a pickup location. We have now added this point to the text.

\item Reviewer 1 wrote:
\begin{quote}
``p. 8, L.25-27: The reason why the fact that the preferences depending on the travel times should reduce the impact of including location specific service rates is not obvious and shuould be expanded on.''
\end{quote}
We have now added a sentence explaining this.

\item Reviewer 1 wrote:
\begin{quote}
``Algorithm 1: Even though it is obvious that the solutions are very likely ranked by decreasing expected survival probability it should be stated clearly, how the population is ranked according to Equation 5.''
\end{quote}
Fixed.

\item Reviewer 1 wrote:
\begin{quote}
``Table 4: The optimised solution shows (slightly) higher mean response times for scenarios D13 and D19. This should be discussed.''
\end{quote}
This is discussed directly in the paragraph discussing the table: ``It is worth noting that for the simulation based KPIs (mean response times and percent abandoned), there is not too much difference between the currently used allocation and the better allocation, showing that these KPIs may not be good approximations of survival.''

\item Reviewer 1 wrote:
\begin{quote}
``The relaxation of the assumption of location independent service rates and and time-independent arrivale rates in the MESLMHPHF could be remarked upon in the outlook''
\end{quote}
This has now been added to the discussion section.

\item Reviewer 1 wrote:
\begin{quote}
``Reframing subchapter 6.3 as e.g. demand scenario definition would help the reading flow''
\end{quote}

\end{itemize}

\subsection{Comments from Reviewer 2}

\begin{itemize}

\item Reviewer 2 wrote:
\begin{quote}
``The introduction makes a case for using survival vs reponse time as a measure for ambulance allocation (although this is not the only measure in the results). What would this do to fairness in the allocations? Would far away patients with serious conditions (ie. Needing a fast response time) just be discarded as it is expensive to cater for them (in terms of overall survival of the population)?''
\end{quote}

\item Reviewer 2 wrote:
\begin{quote}
``Is the survival function the same for $K_a$ patients regardless of which vehicle visits them? What is the rationale for this? It seems that primary vehicles would be better prepared to deal with complications, etc.''
\end{quote}

\item Reviewer 2 wrote:
\begin{quote}
``Have you given any consideration to using an exact method to solve the optimisation problem solved in Algorithm 1? Is there any alternative in the literature?''
\end{quote}

\item Reviewer 2 wrote:
\begin{quote}
``P11 L43 – What is the rationale of using a triangular distribution for travel times? In general, what is the basis of all the travel time assumptions presented in this section?''
\end{quote}

\item Reviewer 2 wrote:
\begin{quote}
``P12 5.4 – For patients that might not need to travel to a hospital, would they still need a primary vehicle after a secondary vehicle tends to them? If not the logic would not be separate entirely.''
\end{quote}

\item Reviewer 2 wrote:
\begin{quote}
``It seems there are two conflicting methods here, the optmisiation with the EA, which uses some averages, etc. And a more comprehensive DES simulation; however the latter is not used for the optimisation. Did you consider simulation optimisation or simheuristics to address this?''
\end{quote}

\item Reviewer 2 wrote:
\begin{quote}
``P15 Paragraph starting L36. Can you give more detail on the modelling of the call arrival rates? For rare specialties, and given the low usage of ambulances in Indonesia, you might not have enough data to give an accurate representation of actual demand. Especially, for future demand where visibility is increased, how will this change for example among richer/poorer parts of the city? I acknowledge it is impossible to include everything in your experiments, but how do you expect these limitations might affect your results?''
\end{quote}

\item Reviewer 2 wrote:
\begin{quote}
``Could you conduct any validation of your results with real data from eg. Ambulance response times with the baseline allocation?''
\end{quote}

\item Reviewer 2 wrote:
\begin{quote}
``P4 L34 – Note that IF .... it is possible''
\end{quote}

\item Reviewer 2 wrote:
\begin{quote}
``P4 L41-42 – Travel back to ITS (not it’s)''
\end{quote}

\item Reviewer 2 wrote:
\begin{quote}
``P7 L29 – equations in APPENDIX A.''
\end{quote}

\item Reviewer 2 wrote:
\begin{quote}
``P8 Fig 3 – Include y axis scale''
\end{quote}

\item Reviewer 2 wrote:
\begin{quote}
``P8 L27 – change ending to – This might not have a strong impact on the results of this study.''
\end{quote}

\item Reviewer 2 wrote:
\begin{quote}
``P8 L49 – In addition to the package used, can you more explicitly state what optimisation problem was solved here, and by which method?''
\end{quote}

\item Reviewer 2 wrote:
\begin{quote}
``P12 L26,27 Rewrite, possible a comma is missing there, also L32.''
\end{quote}

\end{itemize}

\end{document}

### 2024-07-26 response/response.tex

\subsection{Comments from Reviewer 2}

\begin{itemize}

\item Reviewer 2 wrote:
\begin{quote}
``The introduction makes a case for using survival vs reponse time as a measure for ambulance allocation (although this is not the only measure in the results). What would this do to fairness in the allocations? Would far away patients with serious conditions (ie. Needing a fast response time) just be discarded as it is expensive to cater for them (in terms of overall survival of the population)?''
\end{quote}

\item Reviewer 2 wrote:
\begin{quote}
``Is the survival function the same for $K_a$ patients regardless of which vehicle visits them? What is the rationale for this? It seems that primary vehicles would be better prepared to deal with complications, etc.''
\end{quote}
This is an interesting suggestion. Currently we model the survival function to be independent of the vehicle type, that is $s_k$ is the same regardless if the vehicle that arrived is a primary or secondary vehicle. As you note, this may not be realistic, and we may want to be able to model situations where there are different survival functions for each vehicle type. However, this becomes more complicated when both vehicles arrive, but at different times - there will be an interaction between the functions that requires more investigation and is beyond the scope of this study. These scenarios are interesting, and we have added a sentence in the paper explaining this as further work.

@@

\item Reviewer 2 wrote:
\begin{quote}
``It seems there are two conflicting methods here, the optmisiation with the EA, which uses some averages, etc. And a more comprehensive DES simulation; however the latter is not used for the optimisation. Did you consider simulation optimisation or simheuristics to address this?''
\end{quote}

\item Reviewer 2 wrote:
\begin{quote}
``P15 Paragraph starting L36. Can you give more detail on the modelling of the call arrival rates? For rare specialties, and given the low usage of ambulances in Indonesia, you might not have enough data to give an accurate representation of actual demand. Especially, for future demand where visibility is increased, how will this change for example among richer/poorer parts of the city? I acknowledge it is impossible to include everything in your experiments, but how do you expect these limitations might affect your results?''
\end{quote}

\item Reviewer 2 wrote:
\begin{quote}
``Could you conduct any validation of your results with real data from eg. Ambulance response times with the baseline allocation?''
\end{quote}

\item Reviewer 2 wrote:
\begin{quote}
``P4 L34 – Note that IF .... it is possible''
\end{quote}
I believe the fix should be ``Note that having .... is possible''.

\item Reviewer 2 wrote:
\begin{quote}
``P4 L41-42 – Travel back to ITS (not it’s)''
\end{quote}
Fixed.

\item Reviewer 2 wrote:
\begin{quote}
``P7 L29 – equations in APPENDIX A.''
\end{quote}
Fixed throughout.

@@

\item Reviewer 2 wrote:
\begin{quote}
``P8 Fig 3 – Include y axis scale''
\end{quote}

\item Reviewer 2 wrote:
\begin{quote}
``P8 L27 – change ending to – This might not have a strong impact on the results of this study.''
\end{quote}
Fixed.

\item Reviewer 2 wrote:
\begin{quote}
``P8 L49 – In addition to the package used, can you more explicitly state what optimisation problem was solved here, and by which method?''
\end{quote}
Section 4.4 explains this, it is not an optimisation problem. We solve equations 13 and 14, that is we find values of $\lambda_a$ and $\tilde{\lambda}_a$ such that those equations are satisfied. These equations are non-linear and so the values are found using the MINPACK hybrd and hybrj algorithms, implemented in Scipy.

\item Reviewer 2 wrote:
\begin{quote}
``P12 L26,27 Rewrite, possible a comma is missing there, also L32.''
\end{quote}
Fixed.

### 2024-07-26 tex/main.tex
address some comments by reviewer 2

\begin{itemize}
  \item $\lambda_{pk}$ be the rate at which patients of speciality $k$ make
        calls from pick-up location $p$;
  \item $s_k(t)$ be the survival function function associated with patients of
        speciality $k$;
  \item $\pi_a$ be the average utilisation of primary vehicles stationed at
        location $a$;
  \item $b_{pa}$ be the expected travel time from ambulance location $a$ to
        pick-up location $p$; and let $B_{pa}$ be a random variable representing
        the time it takes for an ambulance to drive this distance;
  \item $C_{py}$ be a random variable representing the travel time from pick-up
        location $p$ to hospital $y$;
  \item $D_{ya}$ be a random variable representing the travel time from hospital
        $y$ to ambulance location $a$;
  \item $F_{pa}$ be a random variable representing the travel time from pick-up
        location $p$ to ambulance location $a$;
  \item $G_k$ be a random variable representing the time the ambulance spends
        with patients of speciality $k$ at the pick-up location;
  \item $J_k$ be a random variable representing the time the ambulance spends
        with patients of speciality $k$ at the hospital;
  \item $\Theta$ be a random variable representing the time the ambulance spends
        re-fuelling, re-stocking, and resting between transit jobs;
  \item $q_{pky}$ be the probability that a patient of speciality $k$ from
        pick-up location $p$ is taken to hospital $y$.
        Note that $\sum_{y \in \mathcal{Y}} q_{pky} < 1$ is possible, that is a
        patient may not go to any hospital, in which case the ambulance returns
        to their ambulance location.
\end{itemize}

@@

Patients will call an ambulance from one of the pick-up locations
$p \in \mathcal{P}$. If \textit{all} ambulances are busy, then that call will be
abandoned and it is assumed that the patient will find their own way to the
hospital through private or public transportation, which is an appropriate and
justified assumption for the case of Jakarta \citep{BriceSyaribahNoor2022Esui}.
Otherwise, a central control centre will dispatch the closest available primary
vehicle. That vehicle will travel from its location to the patient ($B_{pa}$),
spend some time on the scene treating the patient ($G_k$), travel to the
hospital ($C_{py}$), spend some time handing over the patient at the hospital
($J_k$), travel back to it's original location ($D_{ya}$), then spend time
refilling and refuelling ($\Theta$).
There is also a probability that the patient does not need a hospital, and so
the vehicle will travel from the pick-up location back to it's original vehicle
location ($F_{pa}$). These routes are shown visually in
Figure~\ref{fig:travel_routes}. Note that these routes cannot be interrupted,
that is, an ambulance must return to it's station for refilling before it is
available to respond to another call, to ensure it is cleaned, and fully stocked
for it's next call. Note also that $B_{pa}$ and $F_{pa}$ need not necessarily be
equal, as travel times need not necessarily be symmetrical, this could be due to
a number of reasons, such as non-symmetric road networks, and differences in
driver urgency on each leg of the journey.

@@

\subsection{Maximal Expected Survival Location Model with Heterogeneous
            Patients and Heterogeneous Fleet}\label{sec:meslmhphf}
Here we propose a Maximal Expected Survival Location Model with Heterogeneous
Patients and Heterogeneous Fleet (MESLMHPHF). It is an extension of the
MESLMHP model given in \citet{Knight2012918}, which did not consider
heterogeneous fleets. This is a model of survival that can be used as an
objective function for optimisation algorithms. It is a weighted expected
survival function as the objective function which considers heterogeneous
patients (different specialities) and heterogeneous fleets (both primary
vehicles, EAs, and secondary vehicles, RRVs), and is constructed by
appropriately summing these survival curves, described in
Section~\ref{sec:survival}, across the population, multiplying by ambulance
availability where appropriate.
\textcolor{red}{The full MESLMHPHF optimisation model is given in
Appendix~\ref{apx:fullmodel}, and its components are described fully here.}
The MESLMHPHF is given by Equation~\ref{eqn:meslmhphf}.

@@

Equation~\ref{eqn:survival_A} is the probability of a patient surviving (their
survival function), multiplied by the availability of primary vehicles at that
ambulance station, multiplied by the unavailability of vehicles at closer
stations.
Equation~\ref{eqn:survival_B} extends this to two vehicles types, the first
part of the sum repeating the logic for the faster secondary vehicles, and the
second part of the sum adapting that logic for primary vehicles, who will only
be contribute the patient's survival if all faster secondary vehicles are also
busy. These interpretations are given as annotations to the equations
in~\ref{apx:annotated}.

@@

\noindent
then, using the simulation described in Section~\ref{sec:simulation}, we
consider the distribution of $\rho$ across all simulated jobs under the current
situation and allocation in Jakarta, shown in Figure~\ref{fig:rho_distribution}.
We see for primary vehicle jobs on average 24\% of a job time is location
dependent, while it is 36.5\% for secondary vehicles. Considering
location-dependent service rates would be an interesting further research
question, potentially improving the model's performance by incorporating travel
times into the MESLMHPHF. However, given that in the case of Jakarta the
preferences $\beta_{p a_1 a_2}$ and $R_{p a_1 a_2}$ are themselves defined by
travel times (Equations~\ref{eqn:beta} and~\ref{eqn:R}), this may not have much
affect for this study, \textcolor{red}{as the model already gives preference to
the locations with the shortest service times.}

@@

This is possible as there are no synchronicity issues between the two
components. Primary vehicles operate independently of secondary vehicles, that
is the way primary vehicles respond to a patient is unaffected by the presence
or lack of a secondary vehicle.
Therefore, primary vehicle logic is not compromised by simulating primary
vehicles in isolation.
Secondary vehicles are impacted by the behaviour of primary vehicles, they
must remain with the patient until the primary vehicle arrives, and so must be
simulated after the simulation of primary vehicles, taking as inputs the exact
list of events that occurred. That is the logic of the secondary vehicles is
determined by observing the actions of primary vehicles and reacting to them.
Simulation results are combined for each individual patient, to determine
their response time caused by either primary or secondary vehicles.

@@

All models are parameterised using demand data that covers all 261
neighbourhoods from 1 January to 31 December 2019, before the COVID-19
pandemic. Data from 2020-2021 was naturally heavily skewed by the pandemic,
with significantly lower demand, and so was not considered as representative
of a typical period for forecasting future needs, hence the decision to use
the available 2019 demand. \ref{apx:parameters} provides further details on the
parameterisation of the model, including how travel time estimates were
obtained.

@@

\subsection{Improving the Current Allocation}\label{sec:optimise_current}
Applying the optimisation metaheuristic from Section~\ref{sec:optimising} to the
current allocation re-allocates the 81 EAs and 13 RRVs. We do this for each of
the four scenarios described in Section~\ref{sec:demand_scenarios}. The
hyper-parameters used for this optimisation algorithm throughout this study are
given in Table~\ref{tbl:hyperparameters}, which include approximations for the
average service times of primary and secondary vehicles derived from the initial
simulation model. \ref{apx:hyperparameters} shows some initial explorations on
the hyper-parameter choices, giving confidence that the chosen parameters, with
a large enough number of iterations $N$, are sufficient to find allocations
that perform well.

>>>>

\begin{itemize}
  \item $\lambda_{pk}$ be the rate at which patients of speciality $k$ make
        calls from pick-up location $p$;
  \item $s_k(t)$ be the survival function function associated with patients of
        speciality $k$;
  \item $\pi_a$ be the average utilisation of primary vehicles stationed at
        location $a$;
  \item $b_{pa}$ be the expected travel time from ambulance location $a$ to
        pick-up location $p$; and let $B_{pa}$ be a random variable representing
        the time it takes for an ambulance to drive this distance;
  \item $C_{py}$ be a random variable representing the travel time from pick-up
        location $p$ to hospital $y$;
  \item $D_{ya}$ be a random variable representing the travel time from hospital
        $y$ to ambulance location $a$;
  \item $F_{pa}$ be a random variable representing the travel time from pick-up
        location $p$ to ambulance location $a$;
  \item $G_k$ be a random variable representing the time the ambulance spends
        with patients of speciality $k$ at the pick-up location;
  \item $J_k$ be a random variable representing the time the ambulance spends
        with patients of speciality $k$ at the hospital;
  \item $\Theta$ be a random variable representing the time the ambulance spends
        re-fuelling, re-stocking, and resting between transit jobs;
  \item $q_{pky}$ be the probability that a patient of speciality $k$ from
        pick-up location $p$ is taken to hospital $y$.
        Note that \textcolor{red}{having} $\sum_{y \in \mathcal{Y}} q_{pky} < 1$ is possible, that is a
        patient may not go to any hospital, in which case the ambulance returns
        to their ambulance location.
\end{itemize}

@@

Patients will call an ambulance from one of the pick-up locations
$p \in \mathcal{P}$. If \textit{all} ambulances are busy, then that call will be
abandoned and it is assumed that the patient will find their own way to the
hospital through private or public transportation, which is an appropriate and
justified assumption for the case of Jakarta \citep{BriceSyaribahNoor2022Esui}.
Otherwise, a central control centre will dispatch the closest available primary
vehicle. That vehicle will travel from its location to the patient ($B_{pa}$),
spend some time on the scene treating the patient ($G_k$), travel to the
hospital ($C_{py}$), spend some time handing over the patient at the hospital
($J_k$), travel back to its original location ($D_{ya}$), then spend time
refilling and refuelling ($\Theta$).
There is also a probability that the patient does not need a hospital, and so
the vehicle will travel from the pick-up location back to it's original vehicle
location ($F_{pa}$). These routes are shown visually in
Figure~\ref{fig:travel_routes}. Note that these routes cannot be interrupted,
that is, an ambulance must return to it's station for refilling before it is
available to respond to another call, to ensure it is cleaned, and fully stocked
for it's next call. Note also that $B_{pa}$ and $F_{pa}$ need not necessarily be
equal, as travel times need not necessarily be symmetrical, this could be due to
a number of reasons, such as non-symmetric road networks, and differences in
driver urgency on each leg of the journey.

@@

\subsection{Maximal Expected Survival Location Model with Heterogeneous
            Patients and Heterogeneous Fleet}\label{sec:meslmhphf}
Here we propose a Maximal Expected Survival Location Model with Heterogeneous
Patients and Heterogeneous Fleet (MESLMHPHF). It is an extension of the
MESLMHP model given in \citet{Knight2012918}, which did not consider
heterogeneous fleets. This is a model of survival that can be used as an
objective function for optimisation algorithms. It is a weighted expected
survival function as the objective function which considers heterogeneous
patients (different specialities) and heterogeneous fleets (both primary
vehicles, EAs, and secondary vehicles, RRVs), and is constructed by
appropriately summing these survival curves, described in
Section~\ref{sec:survival}, across the population, multiplying by ambulance
availability where appropriate.
\textcolor{red}{The full MESLMHPHF optimisation model is given in
APPENDIX~\ref{apx:fullmodel}, and its components are described fully here.}
The MESLMHPHF is given by Equation~\ref{eqn:meslmhphf}.

@@

Equation~\ref{eqn:survival_A} is the probability of a patient surviving (their
survival function), multiplied by the availability of primary vehicles at that
ambulance station, multiplied by the unavailability of vehicles at closer
stations.
Equation~\ref{eqn:survival_B} extends this to two vehicles types, the first
part of the sum repeating the logic for the faster secondary vehicles, and the
second part of the sum adapting that logic for primary vehicles, who will only
be contribute the patient's survival if all faster secondary vehicles are also
busy. These interpretations are given as annotations to the equations
in~\textcolor{red}{APPENDIX}\ref{apx:annotated}.

\textcolor{red}{Note that in Equations~\ref{eqn:survival_A}
and~\ref{eqn:survival_B}, the survival functions used, $s_k(t)$, are the same
whether a primary or a secondary vehicle arrives first. This may be unrealistic,
where primary vehicles may be better equipped than secondary vehicles to deal
with some emergencies. One option to overcome this would be to have distinct
survival functions for each vehicle type, however there may be some interactions
between the survival probability functions that would not be captured, for
example when a primary vehicle arrives after a secondary vehicle, there may
still be a better chance of survival that the secondary vehicle only. This would
be an interesting avenue of further research. For this work, survival functions
will be the same for all vehicles, representing the probability of surviving
given time to be seen by any emergency vehicle.}

@@

\noindent
then, using the simulation described in Section~\ref{sec:simulation}, we
consider the distribution of $\rho$ across all simulated jobs under the current
situation and allocation in Jakarta, shown in Figure~\ref{fig:rho_distribution}.
We see for primary vehicle jobs on average 24\% of a job time is location
dependent, while it is 36.5\% for secondary vehicles. Considering
location-dependent service rates would be an interesting further research
question, potentially improving the model's performance by incorporating travel
times into the MESLMHPHF. However, given that in the case of Jakarta the
preferences $\beta_{p a_1 a_2}$ and $R_{p a_1 a_2}$ are themselves defined by
travel times (Equations~\ref{eqn:beta} and~\ref{eqn:R}), this \textcolor{red}{might not have a strong impact on the results of \st{may not have much
affect for}} this study, \textcolor{red}{as the model already gives preference to
the locations with the shortest service times.}

@@

This is possible as there are no synchronicity issues between the two
components. Primary vehicles operate independently of secondary vehicles, that
is\textcolor{red}{,} the way primary vehicles respond to a patient is unaffected by the presence
or lack of a secondary vehicle.
Therefore, primary vehicle logic is not compromised by simulating primary
vehicles in isolation.
Secondary vehicles are impacted by the behaviour of primary vehicles, they
must remain with the patient until the primary vehicle arrives, and so must be
simulated after the simulation of primary vehicles, taking as inputs the exact
list of events that occurred. That is\textcolor{red}{,} the logic of the secondary vehicles is
determined by observing the actions of primary vehicles and reacting to them.
Simulation results are combined for each individual patient, to determine
their response time caused by either primary or secondary vehicles.

@@

All models are parameterised using demand data that covers all 261
neighbourhoods from 1 January to 31 December 2019, before the COVID-19
pandemic. Data from 2020-2021 was naturally heavily skewed by the pandemic,
with significantly lower demand, and so was not considered as representative
of a typical period for forecasting future needs, hence the decision to use
the available 2019 demand. \textcolor{red}{APPENDIX}~\ref{apx:parameters} provides further details on the
parameterisation of the model, including how travel time estimates were
obtained.

@@

\subsection{Improving the Current Allocation}\label{sec:optimise_current}
Applying the optimisation metaheuristic from Section~\ref{sec:optimising} to the
current allocation re-allocates the 81 EAs and 13 RRVs. We do this for each of
the four scenarios described in Section~\ref{sec:demand_scenarios}. The
hyper-parameters used for this optimisation algorithm throughout this study are
given in Table~\ref{tbl:hyperparameters}, which include approximations for the
average service times of primary and secondary vehicles derived from the initial
simulation model. \textcolor{red}{APPENDIX}~\ref{apx:hyperparameters} shows some initial explorations on
the hyper-parameter choices, giving confidence that the chosen parameters, with
a large enough number of iterations $N$, are sufficient to find allocations
that perform well.
